---
title: "Comparing Signature matrices"
output: html_document
date: "2023-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(SpatialDecon)
```


# Reading all the signature matrices
```{r}
lm22= read_delim("Data/Signature_matrices/LM22_signature_matrix.txt")
til10 = read_xlsx("Data/signature_matrices/TIL10_signature_matrix.xlsx", sheet = "Signature") %>% 
        mutate_at(vars(-GeneSymbol), as.numeric)
immunostates= read_xlsx("Data/Signature_matrices/Immunostates_blood_cell_matrix.xlsx") %>% 
              janitor::row_to_names(2) %>% 
              mutate_at(vars(-Gene), as.numeric)
safetme= safeTME %>% as.data.frame() %>% rownames_to_column(var = "Gene_symbol")
glioma_sce= read_csv("Data/Signature_matrices/Custom_signature_matrix_from_glioma_sce.csv")
tonsil_sce= read_csv("Data/Signature_matrices/Custom_signature_matrix_from_tonsil_sce.csv")

```

# Violoin plots of the expression values of B cells in the different signature matrices
```{r}

background_of_genes_in_signature_matrices = full_join(
lm22 %>% 
  dplyr::rename("Gene_symbol" = "Gene symbol") %>% 
  mutate("B_cell.LM22"= `B cells naive`+`B cells memory`) %>% 
  select(Gene_symbol,B_cell.LM22),

til10 %>% 
  dplyr::rename("Gene_symbol"= "GeneSymbol",
                 "B_cell.TIL10"= "B.cells") %>% 
  select(Gene_symbol,B_cell.TIL10),

by=c("Gene_symbol"="Gene_symbol")) %>% 
full_join(
immunostates %>% 
   dplyr::rename("Gene_symbol"= "Gene") %>% 
   mutate(B_cell.Immunostates= memory_B_cell + naive_B_cell) %>% 
   select(Gene_symbol, B_cell.Immunostates),
by=c("Gene_symbol"="Gene_symbol")) %>% 
full_join(
safetme %>% 
  mutate(B_cell.safeTME= B.naive+B.memory) %>% 
  select(Gene_symbol, B_cell.safeTME),
by=c("Gene_symbol"="Gene_symbol")) %>% 
pivot_longer(cols = starts_with("B_cell"),values_to = "Value", names_to = "Origin") %>% 
mutate(Origin= str_replace(Origin,"B_cell.","")) %>% 
ggplot(aes(x=Origin,y=log(Value), color=Origin))+
  geom_violin() +
  geom_point()

```

# Density distribution of the B cell exxpression values in signature matrices
```{r}
density_distribution_of_B_cell_values = full_join(
lm22 %>% 
  dplyr::rename("Gene_symbol" = "Gene symbol") %>% 
  mutate("B_cell.LM22"= `B cells naive`+`B cells memory`) %>% 
  select(Gene_symbol,B_cell.LM22),

til10 %>% 
  dplyr::rename("Gene_symbol"= "GeneSymbol",
                 "B_cell.TIL10"= "B.cells") %>% 
  select(Gene_symbol,B_cell.TIL10),

by=c("Gene_symbol"="Gene_symbol")) %>% 
full_join(
immunostates %>% 
   dplyr::rename("Gene_symbol"= "Gene") %>% 
   mutate(B_cell.Immunostates= memory_B_cell + naive_B_cell) %>% 
   select(Gene_symbol, B_cell.Immunostates),
by=c("Gene_symbol"="Gene_symbol")) %>% 
full_join(
safetme %>% 
  mutate(B_cell.safeTME= B.naive+B.memory) %>% 
  select(Gene_symbol, B_cell.safeTME),
by=c("Gene_symbol"="Gene_symbol")) %>% 
pivot_longer(cols = starts_with("B_cell"),values_to = "Value", names_to = "Origin") %>% 
mutate(Origin= str_replace(Origin,"B_cell.","")) %>% 
ggplot( aes(x=log(Value), fill=Origin)) + geom_density(alpha=.3)

```

################################################################################
# Testing signature matrices with pseudobulk samples
```{r}
pbmc6k <- readRDS("Raw_annotated_data/pbmc6k.rds")
cell_count_pbmc6k <- as.data.frame(table(pbmc6k$cell))

pbmc8k <- readRDS("Raw_annotated_data/pbmc8k.rds")
cell_count_pbmc8k <- as.data.frame(table(pbmc8k$cell))

liver <- readRDS("Raw_annotated_data/liver.rds")
cell_count_liver <- as.data.frame(table(liver$cell))

signature_reference <- readRDS("Data/signature_reference")
```

```{r}
source("library_loading.R")
source("processing_sce.R")
source("Generating_a_pseudobulk_sample.R")
source("updated_spatial_deconv.R")
source("signature_matrix_bootstrapping_without_subsampling.R")


final_table <-data.frame()

for (i in 1:30)

{
# PBMC6k Dataset
pbmc6k_cell_table <-data.frame("cell_type"= c ("B cells","Monocytes","NK cells","T cells")) %>% 
                    left_join(cell_count_pbmc6k, by = c("cell_type" = "Var1")) %>% rowwise() %>% 
                    mutate(inDeconv = as.integer(runif(1,min=10,max = Freq))) %>%  as.data.frame()

pbmc6k_cell_table <- pbmc6k_cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))

pbmc6k_for_loop<- generating_matrix_for_deconv(pbmc6k,"pbmc6k",pbmc6k_cell_table[1,1],pbmc6k_cell_table[1,3],
                                      pbmc6k_cell_table[2,1],pbmc6k_cell_table[2,3],
                                      pbmc6k_cell_table[3,1],pbmc6k_cell_table[3,3],
                                      pbmc6k_cell_table[4,1],pbmc6k_cell_table[4,3])

## PBMC8K Dataset

pbmc8k_cell_table <-data.frame("cell_type"= c ("B cells","Monocytes","NK cells","T cells")) %>% 
                    left_join(cell_count_pbmc8k, by = c("cell_type" = "Var1")) %>% rowwise() %>% 
                    mutate(inDeconv = as.integer(runif(1,min=10,max=Freq))) %>% as.data.frame()

pbmc8k_cell_table <- pbmc8k_cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))

pbmc8k_for_loop<- generating_matrix_for_deconv(pbmc8k,"pbmc8k",pbmc8k_cell_table[1,1],pbmc8k_cell_table[1,3],
                                      pbmc8k_cell_table[2,1],pbmc8k_cell_table[2,3],
                                      pbmc8k_cell_table[3,1],pbmc8k_cell_table[3,3],
                                      pbmc8k_cell_table[4,1],pbmc8k_cell_table[4,3])

## Immune liver Dataset
liver_cell_table <-data.frame("cell_type"= c ("B cells","Monocytes","NK cells","T cells")) %>% 
                    left_join(cell_count_liver, by = c("cell_type" = "Var1")) %>% rowwise() %>% 
                    mutate(inDeconv = as.integer(runif(1,min=10,max=Freq))) %>% as.data.frame()
liver_cell_table <- liver_cell_table %>% mutate(inDeconv_Frac = inDeconv/sum(inDeconv))

liver_for_loop<- generating_matrix_for_deconv(liver,"liver",liver_cell_table[1,1],liver_cell_table[1,3],
                                      liver_cell_table[2,1],liver_cell_table[2,3],
                                      liver_cell_table[3,1],liver_cell_table[3,3],
                                      liver_cell_table[4,1],liver_cell_table[4,3])

## Merging to make a pseudobulk dataset
deconv_table<- merge(liver_for_loop,pbmc8k_for_loop,by.x="HUGO",by.y="HUGO", all=TRUE) %>% merge(pbmc6k_for_loop,by="HUGO",all=TRUE)

#result <-signature_matrix_bootstrapping_without_subsampling(deconv_table,signature_reference)
result <-signature_matrix_bootstrapping_without_subsampling_without_cell_numbers(deconv_table,signature_reference)


result$bootstrap_run=i

final_table <- rbind(final_table,result) 

print(paste0(i,"/30 of the Bootstrap done"))

rm(pbmc6k_for_loop,pbmc8k_for_loop,liver_for_loop,deconv_table)

}
```

## Data wrangling for plotting
```{r}

cell_deconvolutions <- final_table

pb6k<-cell_deconvolutions[, !names(cell_deconvolutions) %in% c("pbmc8k_TPM","liver_TPM","pbmc8k_actual","liver_actual")] %>% mutate("Dataset"="pbmc6k") %>% dplyr::rename("Predicted_values"="pbmc6k_TPM","True_values"="pbmc6k_actual")

pb8k <- cell_deconvolutions[,!names(cell_deconvolutions) %in% c("pbmc6k_TPM","liver_TPM","pbmc6k_actual","liver_actual")] %>% mutate("Dataset"="pbmc8k") %>% dplyr::rename("Predicted_values"="pbmc8k_TPM","True_values"="pbmc8k_actual")

liv <- cell_deconvolutions[,!names(cell_deconvolutions) %in% c("pbmc8k_TPM","pbmc6k_TPM","pbmc8k_actual","pbmc6k_actual")] %>% mutate("Dataset"="Liver") %>% dplyr::rename("Predicted_values"="liver_TPM","True_values"="liver_actual")

cell_deconvolutions_for_plotting=rbind(pb6k,pb8k,liv)

cell_deconvolutions_for_plotting = cell_deconvolutions_for_plotting %>% mutate(cell_type= ifelse(cell_type=="T cells regulatory (Tregs)","T cell", cell_type))


#write_csv(cell_deconvolutions_for_plotting,file = "Data/Bootstrap_datasignature_matrix_comparison_bootstrapped.csv")
#write_csv(cell_deconvolutions_for_plotting,file = "Data/Bootstrap_datasignature_matrix_comparison_bootstrapped_without_cell_number.csv")

cell_deconvolutions_for_plotting=read_csv("Data/Bootstrap_datasignature_matrix_comparison_bootstrapped.csv")
cell_deconvolutions_for_plotting_without_cell_numbers= read_csv("Data/Bootstrap_datasignature_matrix_comparison_bootstrapped_without_cell_number.csv")
```

# Plotting function
```{r}
deconv_type_plots <- function(cell_deconvolutions_for_plotting,deconv_method, plot_title)
{
  nb.cols <- 18
  mycolors <- colorRampPalette(brewer.pal(11, "Set3"))(nb.cols)
  
  cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type==deconv_method) %>% 
  ggplot(aes(x=(True_values),y=(Predicted_values), color,group=cell_type)) + geom_point(aes( color=cell_type), size=3) +
    scale_color_manual (values = c ( 
      "B cell"  =  "#8DD3C7" , 
      "Macrophage"  =  "#FFFFB3",
      "Monocyte"  =  "#BEBADA", 
      "Neutrophil"  =  "#FB8072", 
      "NK cell"  =  "#80B1D3" , 
      "T cell"  =  "#FDB462" ,
      "Myeloid dendritic cell"  =  "#B3DE69" ,      
      "uncharacterized cell"  =  "#FCCDE5",  
      "cytotoxicity score"  =  "#C9B4A7" , 
      "Macrophage/Monocyte"  =  "#D9D9D9",       
      "Endothelial cell"  =  "#BC80BD" ,  
      "Cancer associated fibroblast"  =  "#CCEBC5" ,
      "Progenitor" = "#573F94",
      "Dendritic cell" = "#B3DE69", 
      "Mast cell"="#E6B9A1",
      "Plasma cell"="#E6E6CA",
      "Fibroblast cell"="#CCEBC5")) +
    geom_smooth(method = "lm", formula = y~x, col = "black") +
    scale_shape_manual(values=c(0,1,2,3, 16, 17,11,4,8))  + ggtitle(plot_title) + theme(plot.title = element_text(hjust = 0.5)) +   
    theme(axis.text.y   = element_text(size=14),
          axis.text.x   = element_text(size=14),
          axis.title.y  = element_text(size=14),
          axis.title.x  = element_text(size=14),
          panel.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.border = element_rect(colour = "black", fill=NA, size=1)) +                                                                                                                                      
    xlab("Actual proportions") +
    ylab("Predicted proportions")
}


#unique(cell_deconvolutions_for_plotting$deconv_type)
```

```{r}
library(ggpubr)
library(RColorBrewer)
signature_matrix_comparison_graphs <- ggarrange(
  deconv_type_plots(cell_deconvolutions_for_plotting ,"Custom_pbmc8k", "Custom PBMC 8k"),
  deconv_type_plots(cell_deconvolutions_for_plotting,"Glioma_sce", "Glioma sce"),
  deconv_type_plots(cell_deconvolutions_for_plotting,"Immunostates", "Immunostates"),
  deconv_type_plots(cell_deconvolutions_for_plotting,"LM22", "LM22"),
  deconv_type_plots(cell_deconvolutions_for_plotting,"Spatial_decon_tme", "safeTME"),
  deconv_type_plots(cell_deconvolutions_for_plotting,"TIL10", "TIL10"),
  widths=c(1,1,1,1,1,1), align="h", nrow =1, ncol=6, common.legend = TRUE, legend = "bottom")

signature_matrix_without_cell_numbers_comparison_graphs <- ggarrange(
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers ,"Custom_pbmc8k", "Custom PBMC 8k"),
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers,"Glioma_sce", "Glioma sce"),
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers,"Immunostates", "Immunostates"),
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers,"LM22", "LM22"),
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers,"Spatial_decon_tme", "safeTME"),
  deconv_type_plots(cell_deconvolutions_for_plotting_without_cell_numbers,"TIL10", "TIL10"),
  widths=c(1,1,1,1,1,1), align="h", nrow =1, ncol=6, common.legend = TRUE, legend = "bottom")

```

```{r}
pdf(file = "Results/Bootstrapped_graphs/Comparison_between_signature_matrices_with_spatial_decon.pdf", width = 15, height = 10)
background_of_genes_in_signature_matrices
density_distribution_of_B_cell_values
annotate_figure(signature_matrix_comparison_graphs, top = text_grob("Comparison between the signature matrices when total number of cells are known", 
               color = "black", face = "bold", size = 14))
annotate_figure(signature_matrix_without_cell_numbers_comparison_graphs, top = text_grob("Comparison between the signature matrices when total number of cells are unknown", 
               color = "black", face = "bold", size = 14))
dev.off()
```

# Stastitical test of Significance
```{r}
custom_8k_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Custom_pbmc8k")
res_custom_8k= cor.test( custom_8k_test$Predicted_values , custom_8k_test$True_values, method = "kendall")

glioma_sce_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Glioma_sce")
res_glioma_sce = cor.test( glioma_sce_test$Predicted_values , glioma_sce_test$True_values, method = "kendall")

immunostates_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Immunostates")
res_immunostates= cor.test( immunostates_test$Predicted_values , immunostates_test$True_values, method = "kendall")

lm22_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="LM22")
res_lm22= cor.test( lm22_test$Predicted_values , lm22_test$True_values, method = "kendall")

tme_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Spatial_decon_tme")
res_tme= cor.test( tme_test$Predicted_values , tme_test$True_values, method = "kendall")

til10_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="TIL10")
res_til10= cor.test( til10_test$Predicted_values , til10_test$True_values, method = "kendall")

```

```{r}
stat_table= data.frame("Signature_matrix"= c("Custom_pbmc8k","Glioma_sce","Immunostates","LM22","Spatial_decon_tme","TIL10"),
  "estimates"= c(res_custom_8k$estimate,res_glioma_sce$estimate,res_immunostates$estimate,res_lm22$estimate,res_tme$estimate,res_til10$estimate),
                          "Pval" = c(res_custom_8k$p.value,res_glioma_sce$p.value,res_immunostates$p.value,res_lm22$p.value,res_tme$p.value,res_til10$p.value) )

stat_table
```

```{r}
custom_8k_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Custom_pbmc8k")
res_custom_8k= cor.test( custom_8k_test$Predicted_values , custom_8k_test$True_values, method = "pearson")

glioma_sce_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Glioma_sce")
res_glioma_sce = cor.test( glioma_sce_test$Predicted_values , glioma_sce_test$True_values, method = "pearson")

immunostates_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Immunostates")
res_immunostates= cor.test( immunostates_test$Predicted_values , immunostates_test$True_values, method = "pearson")

lm22_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="LM22")
res_lm22= cor.test( lm22_test$Predicted_values , lm22_test$True_values, method = "pearson")

tme_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="Spatial_decon_tme")
res_tme= cor.test( tme_test$Predicted_values , tme_test$True_values, method = "pearson")

til10_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type=="TIL10")
res_til10= cor.test( til10_test$Predicted_values , til10_test$True_values, method = "pearson")
```
```{r}
stat_table2= data.frame("Signature_matrix"= c("Custom_pbmc8k","Glioma_sce","Immunostates","LM22","Spatial_decon_tme","TIL10"),
  "estimates"= c(res_custom_8k$estimate,res_glioma_sce$estimate,res_immunostates$estimate,res_lm22$estimate,res_tme$estimate,res_til10$estimate),
                          "Pval" = c(res_custom_8k$p.value,res_glioma_sce$p.value,res_immunostates$p.value,res_lm22$p.value,res_tme$p.value,res_til10$p.value) )

stat_table2
```

```{r}
custom_8k_test= cell_deconvolutions_for_plotting %>% dplyr::filter(deconv_type == "Custom_pbmc8k")
shapiro.test(custom_8k_test$True_values)
shapiro.test(custom_8k_test$Predicted_values)
ggqqplot(custom_8k_test$Predicted_values, ylab = "Predicted_values")
ggqqplot(custom_8k_test$True_values, ylab = "True_values")
res_custom_8k= cor.test(custom_8k_test$Predicted_values, custom_8k_test$True_values, method = "kendall")
```

################## Comparing Deconvolution results in Spatial Decon without inputing cell_numbers ##########################
```{r}

```

